---
title: "Diversity-Density.rmd"
author: "Riley Rettig"
date: "12/10/2020"
output: html_document
---

```{r}
#Libraries
library(tidyverse)
library(neonUtilities)
```
First, I'll practice combining subplots with HARV data. From data info:
"Generating a comprehensive list of plant species for an entire plot requires the aggregatoion of tables. Appropriately linking presence of species documented in the 1m2
subplots and published in table div_1m2Data with species documented in 10 and 100m2
subplots published in table div_10m2Data100m2Data_pub requires the correct named locaƟon (siteID and plotID), year, and bout. This sampling event is captured in the field eventID (plotID.boutNumber.year, e.g., STEI_001.1.2020) that can be used to link the tables. If this field is not populated, it can be created by aggregaƟng the fields plotID, boutNumber, and year (subset from endDate)."

This field is not yet populated, so I will have to populate it myself. Then I can find the species richness.  

```{r}
HARV_data <- loadByProduct(dpID = "DP1.10058.001",
                            site = c("HARV"),
                            check.size = F)

#Unlist to Rstudio environment
list2env(HARV_data, .GlobalEnv)

#Need to populate eventID for both 1m and 10m/100m tables
div_1m2Data = mutate(div_1m2Data,
       eventID = paste(plotID, boutNumber, str_sub(endDate, 1, 4), sep = "."))
div_10m2Data100m2Data =  mutate(div_10m2Data100m2Data,
       eventID = paste(plotID, boutNumber, str_sub(endDate, 1, 4), sep = "."))

# Add column for year
div_1m2Data = mutate(div_1m2Data,
                     year = str_sub(endDate, 1, 4))
div_10m2Data100m2Data = mutate(div_10m2Data100m2Data,
                     year = str_sub(endDate, 1, 4))

# Just grab what we need (we'll leave out percentCover for now, even though we'll use it for density later)
# Filter out last two years since they are irregular
species_1m2 <- div_1m2Data %>% 
  filter(divDataType == 'plantSpecies') %>%
  filter(year < 2018) %>%
  select(domainID, siteID, nlcdClass, plotID, endDate, taxonID, scientificName, taxonRank, family, eventID)

species_10m2_100m2 <- div_10m2Data100m2Data %>% 
  filter(year < 2018) %>%
  select(domainID, siteID, nlcdClass, plotID, endDate, taxonID, scientificName, taxonRank, family, eventID)

# Now we can combine with rbind
species <- rbind(species_1m2, species_10m2_100m2)

# Remove duplicate species
uniqueSpecies <- species %>% distinct(species$eventID, species$taxonID, .keep_all = TRUE)

# Now we have to get species count for each unique eventID
species_per_event <- as.data.frame(table(uniqueSpecies$eventID))



# Add back the plotID
species_per_event <- mutate(species_per_event,
                            plotID = str_sub(Var1, 1, 8))

# For now I'll do all of the years, but I wonder if we should cut the most recent year since they seem a lot different, we can see this in the distribution, such as in plot 6
HARV006_species <- species %>% filter(plotID == 'HARV_006')
ggplot(data = HARV006_species, aes(x = endDate)) +
  geom_histogram()

# Find the average for each plot across all years
# FIXME I would like to be able to include other plot info in this data frame
by_plot <- group_by(species_per_event, plotID)
HARV_SR <- as.data.frame(summarise(by_plot, speciesRichness = mean(Freq)))


```
Now that I have the species richness for each plot I will use the sum of percentCover as a measurement for density, although I am not sure whether this is actually a good estimate of density as it is only taken on the 1m2 plots and is done ocularly without getting a count on how many individuals there are. I would need counts of individual plants on these sites which I may be able to get with the woody and herbaceous datasets, but I will look at that later this week.

Note: Percent cover is measured in 6 1m2 subplots, but Prior to 2019: Two addiƟonal 1 and 10m2
subplots were sampled. The presence and percent cover
of species and ancillary data was observed in eight 1m2
subplots. 

Because of this, I will sum up total percent coverage of each subplot then take the average of all subplots for the year, and finally the average of all years

Note on percentages: "estimates should not exceed 100
percent for a single species, but the combined cover of multiple species – even just the biotic
component of the observation - may be greater than 100%""

```{r}
# Select percentCover from data
coverage_1m2 <- div_1m2Data %>% 
  filter(year < 2018) %>%
  filter(divDataType == 'plantSpecies') %>%
  select(domainID, siteID, taxonID,plotID, endDate, subplotID, eventID, percentCover, year)

# Get sum percent cover for each unique subplot by grouping by subplotID
by_subplot <- group_by(coverage_1m2, subplotID, eventID, siteID, plotID, endDate, year)

subplot_coverage <- summarise(by_subplot, totalCoverage = sum(percentCover, na.rm = TRUE))

#Some percentages are over 100%, but this is okay
ggplot(data = subplot_coverage, aes(totalCoverage)) +
  geom_histogram() 
#+
#  coord_cartesian(ylim = c(0, 40), xlim = c(100, 250))

# Now we can get an average for each year
by_year <- group_by(subplot_coverage, plotID)
HARV_density <-  summarise(by_year, density = mean(totalCoverage))
```

Now we can look at the relationship between diversity and density at the HARV site

```{r}
# join the two tables together
HARV_analysis <- left_join(HARV_SR, HARV_density, 
               by=c('plotID'))

# Create a scatterplot
ggplot(HARV_analysis, aes(x = density, y = speciesRichness)) +
  geom_point()+
  geom_smooth(method = "lm")  #linear regression plot

# Pearson's correlation coeficient
# Species Richness is positively correlated with density at HARV (r(31) = 0.71, p < 0.01)
cor.test(HARV_analysis$speciesRichness,
         HARV_analysis$density)

# Linear Regression
# species richness linear regression (b = 0.81, t(31) = 0.494, p < 0.01)
# There was a significant increase in species richness with increasing density
HARVmap <- lm(speciesRichness ~ density, 
                    data = HARV_analysis)
summary(HARVmap)

```

Now we'll scale this up to all sites and see whether the same holds true!

```{r}

#Download plant presence data for all sites where data is available 
PlantPresence_data <- loadByProduct(dpID = "DP1.10058.001", 
                                    check.size = F)
#Unlist to Rstudio environment
list2env(PlantPresence_data, .GlobalEnv)


```




```{r}

#Need to populate eventID for both 1m and 10m/100m tables
div_1m2Data = mutate(div_1m2Data,
       eventID = paste(plotID, boutNumber, str_sub(endDate, 1, 4), sep = "."))
div_10m2Data100m2Data =  mutate(div_10m2Data100m2Data,
       eventID = paste(plotID, boutNumber, str_sub(endDate, 1, 4), sep = "."))

# Add column for year
div_1m2Data = mutate(div_1m2Data,
                     year = str_sub(endDate, 1, 4))
div_10m2Data100m2Data = mutate(div_10m2Data100m2Data,
                     year = str_sub(endDate, 1, 4))

# Just grab what we need (we'll leave out percentCover for now, even though we'll use it for density later)
# Filter out last two years since they are irregular
species_1m2 <- div_1m2Data %>% 
  filter(divDataType == 'plantSpecies') %>%
  filter(year < 2018) %>%
  select(domainID, siteID, nlcdClass, plotID, endDate, taxonID, scientificName, taxonRank, family, eventID)

species_10m2_100m2 <- div_10m2Data100m2Data %>% 
  filter(year < 2018) %>%
  select(domainID, siteID, nlcdClass, plotID, endDate, taxonID, scientificName, taxonRank, family, eventID)

# Now we can combine with rbind
species <- rbind(species_1m2, species_10m2_100m2)

# Remove duplicate species
uniqueSpecies <- species %>% distinct(species$eventID, species$taxonID, .keep_all = TRUE)

# Now we have to get species count for each unique eventID
species_per_event <- as.data.frame(table(uniqueSpecies$eventID))



# Add back the plotID
species_per_event <- mutate(species_per_event,
                            plotID = str_sub(Var1, 1, 8))

# Find the average for each plot across all years
# FIXME I would like to be able to include other plot info in this data frame
by_plot <- group_by(species_per_event, plotID)
SpeciesRichness <- as.data.frame(summarise(by_plot, speciesRichness = mean(Freq)))


```


```{r}
# Select percentCover from data
coverage_1m2 <- div_1m2Data %>% 
  filter(year < 2018) %>%
  filter(divDataType == 'plantSpecies') %>%
  select(domainID, siteID, taxonID,plotID, endDate, subplotID, eventID, percentCover, year)

# Get sum percent cover for each unique subplot by grouping by subplotID
by_subplot <- group_by(coverage_1m2, subplotID, eventID, siteID, plotID, endDate, year)

subplot_coverage <- summarise(by_subplot, totalCoverage = sum(percentCover, na.rm = TRUE))

#Some percentages are again over 100%, but this is okay
ggplot(data = subplot_coverage, aes(totalCoverage)) +
  geom_histogram() 
#+
#  coord_cartesian(ylim = c(0, 40), xlim = c(100, 250))

# Now we can get an average for each year
by_year <- group_by(subplot_coverage, plotID)
Density <-  summarise(by_year, density = mean(totalCoverage))
```

Now we can look at the relationship between diversity and density at all sites, and we find that they are still positively correlated. 

```{r}
# join the two tables together
Analysis <- left_join(SpeciesRichness, Density, 
               by=c('plotID'))

# Create a scatterplot
ggplot(Analysis, aes(x = density, y = speciesRichness)) +
  geom_point()+
  geom_smooth(method = "lm")  #linear regression plot

# Pearson's correlation coeficient
# Species Richness is positively correlated with density at HARV (r(1376) = 0.37, p < 0.01)
cor.test(Analysis$speciesRichness,
         Analysis$density)

# Linear Regression
# species richness linear regression (b = 0.33, t(1376) = 0.1354, p < 0.01)
# There was a significant increase in species richness with increasing density
map <- lm(speciesRichness ~ density, 
                    data = Analysis)
summary(map)

```

Lets join some of the original information together with Analysis and see if we can find any patterns

```{r}

```
